---
title: "Baseball"
author: "Hanying Ji; Jiaqian Yu"
date: "2018/4/19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Setting Environment
```{r,warning=FALSE,message=FALSE}
library(Lahman)
library(dplyr)
library(tidyr)
library(MASS)
library(gamlss)
```

##Data Preparation
Our dataset is from package "Lahman" in R, providing the tables from the "Sean Lahman Baseball Database"
```{r}
# Define pitchers: Those players whose total amounts of pitch > 3
pitchers <-Pitching %>%
  group_by(playerID) %>%
  summarize(gamesPitched=sum(G)) %>%
  filter(gamesPitched>3)

#collect data of batters from 2000 to 2016
info<-Batting %>%
  filter(AB>0,yearID%in%c(2000:2016)) %>%
  anti_join(pitchers, by="playerID") 

info<-info%>%
  dplyr::select(playerID, AB, H) %>%
  group_by(playerID) %>%
  summarize(Hits=sum(H),AtBats=sum(AB))%>%
  mutate(BattingAvg=Hits/AtBats) 

# add batters' names and hand preferance for batting
allinfo<-Master %>%
  tbl_df() %>%
  dplyr::select(playerID, nameFirst, nameLast, bats) %>%
  unite(name, nameFirst, nameLast, sep = " ") %>%
  inner_join(info, by="playerID")
head(allinfo)
```

# Model 1
```{r}
# lowest BattingAvg
head(allinfo[order(allinfo$BattingAvg,decreasing = TRUE),])
# highest BattingAvg
head(allinfo[order(allinfo$BattingAvg,decreasing = FALSE),])
```


# Model 2
```{r}
set.seed(1)
alpha<-seq(1,200,1)
beta<-367*alpha/133
pbeta(0.3,alpha,beta)
```


```{r}
alpha0<-126
beta0<-347.6842
set.seed(1)
df<-data.frame(x=rbeta(1000,alpha0,beta0))
library(ggplot2)
ggplot(df,aes(df$x))+
  stat_function(fun=dbeta,args=list(alpha0,beta0),color="red")+
  xlim(0,0.5)+
  xlab("x")+
  ylab("Density of Beta(119,328)")
allinfo$alpha1<-alpha0+allinfo$Hits
allinfo$beta1<-beta0+allinfo$AtBats-allinfo$Hits
allinfo$BattingAvg_post<-allinfo$alpha1/(allinfo$alpha1+allinfo$beta1)
# lowest BattingAvg_post
head(allinfo[order(allinfo$BattingAvg_post,decreasing = TRUE),])
# highest BattingAvg_post
head(allinfo[order(allinfo$BattingAvg_post,decreasing = FALSE),])
```
```{r}
Top5<- allinfo[order(allinfo$BattingAvg_post,decreasing = TRUE),][1:5,]$playerID
Top5_old<-allinfo[order(allinfo$BattingAvg,decreasing = TRUE),][1:5,]$playerID
# Tail5<-allinfo[order(allinfo$BattingAvg_post,decreasing = FALSE),][1:5,]$playerID
Top5info <- allinfo %>%
   filter(playerID %in% c(Top5,Top5_old))

library(broom)
five <- Top5info %>%
   tidyr::crossing(x = seq(.18, .4, .0002)) %>%
   ungroup() %>%
   mutate(density = dbeta(x, alpha1, beta1))
ggplot(five) +
  geom_line(aes(x, density,color=name)) +
  stat_function(fun = function(x) dbeta(x, alpha0, beta0),
                lty = 2, color = "black")
```



```{r}
ggplot(data=allinfo)+
  geom_point(mapping=aes(BattingAvg,BattingAvg_post,colour=AtBats))+
  scale_colour_gradientn(colours = blues9)+
  geom_hline(yintercept = alpha0/(alpha0+beta0))+
  geom_abline(slope=1,color="blue")
```

There are a lot of method to fit a probability distribution from data in R. We???ll use the fitdistr function from MASS
```{r,warning=FALSE}
# filter players
allinfo_filtered<-allinfo %>% filter(AtBats>=500)
# use the filtered data to fit the empirical Bayes estimation - a Beta prior
mle<-MASS::fitdistr(allinfo_filtered$BattingAvg, dbeta,
                  start = list(shape1 = 1, shape2 = 10))
alpha0_eb<-mle$estimate[1]
beta0_eb<-mle$estimate[2]
miu0_eb<-alpha0_eb/(alpha0_eb+beta0_eb)

ggplot(data=allinfo_filtered)+
  geom_histogram(binwidth = 0.005,aes(x=BattingAvg,y=..density..))+
  stat_function(fun = function(x) dbeta(x, alpha0_eb, beta0_eb), color = "red",size = 1)+
  labs(title="Prior distribution")
```
Then, for each player, they all have their Empericical Bayesian estimators for their BattingAvg based on the prior distribution given by the representative players
```{r}
allinfo<-allinfo%>%
  mutate(EB_BattingAvg_post=(Hits+alpha0_eb)/(AtBats+alpha0_eb+beta0_eb))
```

```{r}
# lowest EB estimator BattingAvg
head(allinfo[order(allinfo$EB_BattingAvg_post,decreasing = TRUE),])
# highest EB estimator BattingAvg
head(allinfo[order(allinfo$EB_BattingAvg_post,decreasing = FALSE),])
```
As we can see, the Empricial Bayes didn't simply choose those players who only have one or two bats, instead, players with large amount of AtBats were chosen.


```{r}
career_eb <- allinfo %>%
  mutate(eb_estimate = (Hits + alpha0_eb) / (AtBats + alpha0_eb + beta0_eb))%>%
  mutate(alpha1_eb = Hits + alpha0_eb,
           beta1_eb = AtBats - Hits + beta0_eb)%>%
  dplyr::select(playerID,name,Hits,AtBats,BattingAvg,eb_estimate,alpha1_eb,beta1_eb)


career_eb <- career_eb %>%
    mutate(low = qbeta(.025, alpha1_eb, beta1_eb),
           high = qbeta(.975, alpha1_eb, beta1_eb))

set.seed(2018)

some <- career_eb %>%
    sample_n(20) %>%
    mutate(name = paste0(name, " (", Hits, "/", AtBats, ")"))

frequentist <- some %>%
    group_by(playerID, name, AtBats) %>%
    do(tidy(binom.test(.$Hits, .$AtBats))) %>%
    dplyr::select(playerID, name, estimate, low = conf.low, high = conf.high) %>%
    mutate(method = "Confidence")

bayesian <- some %>%
    dplyr::select(playerID, name, AtBats, estimate = eb_estimate,
           low = low, high = high) %>%
    mutate(method = "Credible")

combined <- bind_rows(frequentist, bayesian)

combined %>%
    #mutate(name = reorder(name, -AtBats)) %>%
    ggplot(aes(estimate, name, color = method, group = method)) +
    geom_point() +
    geom_errorbarh(aes(xmin = low, xmax = high)) +
    geom_vline(xintercept = alpha0_eb / (alpha0_eb + beta0_eb), color = "red", lty = 2) +
    xlab("Estimated batting average") +
    ylab("Player") +
    labs(color = "")
```


#Model 3
```{r}
library(ggplot2)
allinfo %>%
  filter(AtBats >= 20) %>%
  ggplot(aes(AtBats, BattingAvg)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_log10()
```

```{r}
#fit the initial model
fit <- gamlss(cbind(Hits, AtBats - Hits) ~ log(AtBats),
              data = allinfo,
              family = BB(mu.link = "identity"))
library(broom)
td <- tidy(fit)
td
```
```{r}
#calculate some prior dist
u0 <- 0.1517
lamda <- 0.0143
sigma0 <- 0.0012
u1 <- u0+lamda*log(1)
u2 <- u0+lamda*log(10)
u3 <- u0+lamda*log(100)
u4 <- u0+lamda*log(1000)
u5 <- u0+lamda*log(10000)

a1 <- u1/sigma0
a2 <- u2/sigma0
a3 <- u3/sigma0
a4 <- u4/sigma0
a5 <- u5/sigma0
b1 <- (1-u1)/sigma0
b2 <- (1-u2)/sigma0
b3 <- (1-u3)/sigma0
b4 <- (1-u4)/sigma0
b5 <- (1-u5)/sigma0

df<-data.frame(x=seq(0.1,0.4,1000))
ggplot(df,aes(df$x),col())+
  stat_function(fun = dbeta,args=list(a1,b1), color = "red",size = 1)+
  stat_function(fun = dbeta,args=list(a2,b2), color = "yellow",size = 1)+
  stat_function(fun = dbeta,args=list(a3,b3), color = "green",size = 1)+
  stat_function(fun = dbeta,args=list(a4,b4), color = "blue",size = 1)+
  stat_function(fun = dbeta,args=list(a5,b5), color = "purple",size = 1)+
    xlim(0.1,0.4)+
  labs(title="Updated Prior distribution",x="Batting Average",y="Prior Density")
```



#Model 4
```{r}
#left hand and right hand
career2 <- allinfo %>%
  filter(!is.na(bats)) %>%
  mutate(bats = relevel(bats,"R"))

fit2 <- gamlss(cbind(Hits,AtBats-Hits)~log(AtBats)+bats,
               data=career2,
               family=BB(mu.link="identity"))
tidy(fit2)
```

```{r}
#calculate some prior dist
u0 <- 0.1514
lamda <- 0.0143
phi <- 0.0017
sigma0 <- 0.0012
BattingHand <- c(1,0,1,0,1,0,1,0,1,0)
AB <- c(1,10,100,1000,10000)

u1l <- u0+lamda*log(1)+phi*1
u1r <- u0+lamda*log(1)+phi*0
u2l <- u0+lamda*log(10)+phi*1
u2r <- u0+lamda*log(10)+phi*0
u3l <- u0+lamda*log(100)+phi*1
u3r <- u0+lamda*log(100)+phi*0
u4l <- u0+lamda*log(1000)+phi*1
u4r <- u0+lamda*log(1000)+phi*0
u5l <- u0+lamda*log(10000)+phi*1
u5r <- u0+lamda*log(10000)+phi*0
a1l <- u1l/sigma0
a1r <- u1r/sigma0
a2l <- u2l/sigma0
a2r <- u2r/sigma0
a3l <- u3l/sigma0
a3r <- u3r/sigma0
a4l <- u4l/sigma0
a4r <- u4r/sigma0
a5l <- u5l/sigma0
a5r <- u5r/sigma0



b1l <- (1-u1l)/sigma0
b1r <- (1-u1r)/sigma0
b2l <- (1-u2l)/sigma0
b2r <- (1-u2r)/sigma0
b3l <- (1-u3l)/sigma0
b3r <- (1-u3r)/sigma0
b4l <- (1-u4l)/sigma0
b4r <- (1-u4r)/sigma0
b5l <- (1-u5l)/sigma0
b5r <- (1-u5r)/sigma0

df<-data.frame(x=seq(0.1,0.35,1000))
ggplot(df,aes(df$x))+
  stat_function(fun = dbeta,args=list(a1l,b1l), color = "red",size = 1)+
  stat_function(fun = dbeta,args=list(a2l,b2l), color = "yellow",size = 1)+
  stat_function(fun = dbeta,args=list(a3l,b3l), color = "green",size = 1)+
  stat_function(fun = dbeta,args=list(a4l,b4l), color = "blue",size = 1)+
  stat_function(fun = dbeta,args=list(a5l,b5l), color = "purple",size = 1)+
  stat_function(fun = dbeta,args=list(a1r,b1r), color = "red",size = 1,linetype="dashed")+
  stat_function(fun = dbeta,args=list(a2r,b2r), color = "yellow",size = 1,linetype="dashed")+
  stat_function(fun = dbeta,args=list(a3r,b3r), color = "green",size = 1,linetype="dashed")+
  stat_function(fun = dbeta,args=list(a4r,b4r), color = "blue",size = 1,linetype="dashed")+
  stat_function(fun = dbeta,args=list(a5r,b5r), color = "purple",size = 1,linetype="dashed")+
    xlim(0.1,0.35)+
  labs(title="Updated Prior distribution",x="Batting Average",y="Prior Density")
```
